How to measure the radial velocity of a star:\\The invariant-spectrum case

### 2022-02-03 paper/main.tex

% to-do list
% ----------
% - Look for all instances of HOGG, MB, and CITE.
% - Make an acronym command and fix CLRB and all other acronyms.
% - Be sure to discuss SNR in a continuum resolution element, as well as in a pixel.
%.- Make sure that, for any assumption that is explicitly tested, the forthcoming test is mentioned when the assumption is introduced.
% - Are we clear that we exponentiate the model to get the true data expectation?
% - Are we consistent in our use of Doppler shift, radial velocity, and redshift and so on?
% - Are we sensible in our use of first person?
% - Do we address the relationship between our tiny spectral segment and a full spectrograph? Get scalings. Etc.

@@

In order to test the maximum-likelihood estimator and the estimators to follow, we build a sandbox of completely fake data with known properties, obeying our assumptions.
\figref{fig:data} shows a few example spectra generated from our model.
The model spectra are made with an exactly unit continuum and a set of discrete absorption lines of varying equivalent widths.
The lines are drawn from a uniform Poisson process in the long-wavelength $x$ domain.
Their equivalent widths are drawn from a power-law distribution function.
The spectrograph line-spread function is set to be a Gaussian in the log-wavelength $x$ space, with a standard deviation $\sigma_x=1/R$, where $R=135\,000$ is the spectrograph resolution.
Technically the Gaussian line-spread function is used to generate the logarithm of the flux, which is then exponentiated to deliver the pixel flux expectations. HOGG SHOULD WE HAVE AN EQUATION HERE?
The noise on the continuum-normalized spectral pixel values is Gaussian with zero mean and variance proportional to the continuum-normalized spectral expectation (true intensity).
This noise is explicitly white; it is designed to model photon or shot noise.
The noise variance model is normalized to hit a particular signal-to-noise ratio (SNR) in the continuum (that is, at a continuum-normalized flux $y=1$); typically we set this SNR to 100 (per pixel).
The true Doppler shifts $\delta x\equiv\ln\sqrt{(c+v)/(c-v)}\approx v/c$ for the different spectra are drawn from a uniform distribution in the interval $-1\times10^{-4}<\delta x<1\times10^{-4}$, corresponding to something like barycentric variation.
We included no telluric absorption in these toy data.

@@

\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    [HOGG: NEED FIGURE HERE]
    \end{center}
    \caption{HOGG: Something that shows the maximum-likelihood radial-velocity estimates for all epochs, that they saturate the bound.\label{fig:mlrvs}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/ccf.png}
    \end{center}
    \caption{HOGG: Something that shows the CCF for one epoch, and a zoom in showing the parabola fit.\label{fig:ccf}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/ccfrvs.png}
    \end{center}
    \caption{HOGG: Something that shows the CCF RVs for all epochs, that they saturate the bound.\label{fig:ccfrvs}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/binarymask.png}
    \end{center}
    \caption{HOGG: A visualization of part of the binary mask.\label{fig:binarymask}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/bccfrvswrong.png}
    \end{center}
    \caption{HOGG: Something that shows the binary-mask CCF RVs are terrible if naively optimized.\label{fig:bccfrvswrong}}
  \end{mdframed}
\end{figure}%
\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/bccfrvs.png}
    \end{center}
    \caption{HOGG: Something that shows the binary-mask CCFs deliver good RVs when fit after the fact with a Gaussian function.\label{fig:bccfrvs}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[t]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/empiricalrvs.png}
    \end{center}
    \caption{HOGG: Something that shows the empirical-spectrum CCFs deliver good RVs.\label{fig:empiricalrvs}}
  \end{mdframed}
\end{figure}

>>>>

% to-do list
% ----------
% - Look for all instances of HOGG, MB, and CITE.
% - Make an acronym command and fix CLRB and all other acronyms.
% - Be sure to discuss SNR in a continuum resolution element, as well as in a pixel.
%.- Make sure that, for any assumption that is explicitly tested, the forthcoming test is mentioned when the assumption is introduced.
% - Are we clear that we exponentiate the model to get the true data expectation?
% - Are we consistent in our use of Doppler shift, radial velocity, and redshift and so on?
% - Are we sensible in our use of first person?
% - Do we address the relationship between our tiny spectral segment and a full spectrograph? Get scalings. Etc.
% - Make sure all figures are explicitly referred to in the text.

@@

In order to test the maximum-likelihood estimator and the estimators to follow, we build a sandbox of completely fake data with known properties, obeying our assumptions.
\figref{fig:data} shows a few example spectra generated from our model.
The model spectra are made with an exactly unit continuum and a set of discrete absorption lines of varying equivalent widths.
The lines are drawn from a uniform Poisson process in the long-wavelength $x$ domain.
Their equivalent widths are drawn from a power-law distribution function.
The spectrograph line-spread function is set to be a Gaussian in the log-wavelength $x$ space, with a standard deviation $\sigma_x=1/R$, where $R=135\,000$ is the spectrograph resolution.
Technically the Gaussian line-spread function is used to generate the logarithm of the flux, which is then exponentiated to deliver the pixel flux expectations. HOGG SHOULD WE HAVE AN EQUATION HERE?
The noise on the continuum-normalized spectral pixel values is Gaussian with zero mean and variance proportional to the continuum-normalized spectral expectation (true intensity).
This noise is explicitly white; it is designed to model photon or shot noise.
The noise variance model is normalized to hit a particular signal-to-noise ratio (SNR) in the continuum (that is, at a continuum-normalized flux $y=1$); typically we set this SNR to 100 (per pixel).
The true Doppler shifts $\delta x\equiv\ln\sqrt{(c+v)/(c-v)}\approx v/c$ for the different spectra are generated by a pure sinusoid of the epoch number with an amplitude of $1\times10^{-4}$, corresponding to something like barycentric variation.
We included no telluric absorption in these toy data.

@@

\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    [HOGG: NEED FIGURE HERE]
    \end{center}
    \caption{HOGG: Something that shows the maximum-likelihood radial-velocity estimates for all epochs, that they saturate the bound.\label{fig:mlrvs}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/ccf.png}
    \end{center}
    \caption{HOGG: Something that shows the CCF for one epoch, and a zoom in showing the parabola fit.\label{fig:ccf}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/ccfrvs.png}
    \end{center}
    \caption{HOGG: Something that shows the CCF RVs for all epochs, that they saturate the bound.\label{fig:ccfrvs}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/binarymask.png}
    \end{center}
    \caption{HOGG: A visualization of part of the binary mask.\label{fig:binarymask}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/bccfrvswrong.png}
    \end{center}
    \caption{HOGG: Something that shows the binary-mask CCF RVs are terrible if naively optimized.\label{fig:bccfrvswrong}}
  \end{mdframed}
\end{figure}%
\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/bccfrvs.png}
    \end{center}
    \caption{HOGG: Something that shows the binary-mask CCFs deliver good RVs when fit after the fact with a Gaussian function.\label{fig:bccfrvs}}
  \end{mdframed}
\end{figure}

@@

\begin{figure}[tp]
  \begin{mdframed}
    \begin{center}
    \includegraphics[width=\textwidth]{../notebook/empiricalrvs.png}
    \end{center}
    \caption{HOGG: Something that shows the empirical-spectrum CCFs deliver good RVs.\label{fig:empiricalrvs}}
  \end{mdframed}
\end{figure}
